### 2018-12-16 Draft.Rmd
.

---
title: "Group Fou\\*R* Term Project"
subtitle: "MEF University - BDA 503 Essentials of Data Analytics - Fall 2018"
date: "December 2018"
output: html_document
---
## Explanatory Data Analysis of Trade Statistics of Turkey

### Abstract

We worked on trade statistics data of Turkey against the G7 countries and 3 more countries (Russia, India, China) both for export and import.Our aim is to analyze Turkey?s trade characteristics by product groups. Within this aim, we take trade values and quantities by product groups in MS Excel file format from trademap.org website for each trade partner country for the period of 2013 - 2017. 
In this study, We applied data manipulation to make it proper for analysis and explanatory data analysis with visulations and conclude with overall statistical overview of Turkey's trade.

### Explanation of Raw Data

Following the extraction from source (Trademap.org), our raw data consist of 10 countries? trade data against Turkey. This data splits to 2 with trade direction as ?Export? and ?Import?. The other dimension is stat for trade, one is ?Quantity? of Export/Import, the other is ?Value? in thousand USD of Export/Import. Therefore we have four main data frame for each country (total of 40 excel files).

<table>
<thead>
<tr class="header">
<th align="left">Data Frame</th>
<th align="left">DF Definition</th>
</tr>
</thead>

<tbody>
<tr class="odd">
<td align="left">ExportValues</td>
<td align="left">Export from Turkey to target countries and their values in thousand USD</td>
</tr>
<tr class="odd">
<td align="left">ExportQuantity</td>
<td align="left">Export from Turkey to target countries and their quantities in units(tons,gallons etc)</td>
</tr>
<tr class="odd">
<td align="left">ImportValues</td>
<td align="left">Import to Turkey from target counties and their values in thousand USD</td>
</tr>
<tr class="odd">
<td align="left">ImportQuantity</td>
<td align="left">Import to Turkey from target counties and their quantities in units(tons,gallons etc)</td>
</tr>

</tbody>
</table>

Table structures for "Quantity" and "Value" files are as follows:

<table>
<thead>
<tr class="header">
<th align="left">Country</th>
<th align="left">Product code</th>
<th align="left">Product label</th>
<th align="left">Quantity in 2013</th>
<th align="left">Unit</th>
<th align="left">Quantity in 2014</th>
<th align="left">Unit</th>
<th align="left">Quantity in 2015</th>
<th align="left">Unit</th>
<th align="left">Quantity in 2016</th>
<th align="left">Unit</th>
<th align="left">Quantity in 2017</th>
<th align="left">Unit</th>
</tr>
</thead>

<tbody>
<tr class="odd">
<td align="left">France</td>
<td align="left">'0306</td>
<td align="left">Crustaceans, whether in shell or not, live, fresh, chilled, frozen, dried, salted or in brine, ...</td>
<td align="left">121</td>
<td align="left">Tons</td>
<td align="left">139</td>
<td align="left">Tons</td>
<td align="left">22</td>
<td align="left">Tons</td>
<td align="left">24</td>
<td align="left">Tons</td>

</tr>
</tbody>
</table>

Table 1: Quantity excel table structure

<table>
<thead>
<tr class="header">
<th align="left">Country</th>
<th align="left">Product code</th>
<th align="left">Product label</th>
<th align="left">Value in 2013</th>
<th align="left">Value in 2014</th>
<th align="left">Value in 2015</th>
<th align="left">Value in 2016</th>
<th align="left">Value in 2017</th>
</tr>
</thead>

<tbody>
<tr class="odd">
<td align="left">France</td>
<td align="left">'0306</td>
<td align="left">Crustaceans, whether in shell or not, live, fresh, chilled, frozen, dried, salted or in brine, ...</td>
<td align="left">0</td>
<td align="left">1155</td>
<td align="left">1182</td>
<td align="left">270</td>
<td align="left">0</td>
</tr>
</tbody>
</table>

Table 2: Values excel table structure

In addition to main raw data, we also have look-up table for product codes; This lookup table registers product labels in 2 digit code that is less categorized than main data.

<table>
<thead>
<tr class="header">
<th align="left">Product code</th>
<th align="left">Product label</th>
</tr>
</thead>

<tbody>
<tr class="odd">
<td align="left">'01</td>
<td align="left">Live animals</td>
</tr>
<tr class="odd">
<td align="left">'02</td>
<td align="left">Meat and edible meat offal</td>
</tr>
<tr class="odd">
<td align="left">'03</td>
<td align="left">Fish and crustaceans, molluscs and other aquatic invertebrates</td>
</tr>
<tr class="odd">
<td align="left">'04</td>
<td align="left">Dairy produce; birds' eggs; natural honey; edible products of animal origin, not elsewhere ...</td>
</tr>

</tbody>
</table>

Table 3: Product look-up excel table

Regarding the variables included in these tables; definitions are as follows:

<table>
<thead>
<tr class="header">
<th align="center" colspan="2">ExportQuantity & ImportQuantity</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="left">Country</td>
<td align="left">Trade Partner of Turkey</td>
</tr>
<tr class="odd">
<td align="left">Product_Code</td>
<td align="left">Traded Product Code in 4 digit</td>
</tr>
<tr class="odd">
<td align="left">Product_Label</td>
<td align="left">Product Definition</td>
</tr>  
<tr class="odd">
<td align="left">Unit</td>
<td align="left">Metric of Quantity</td>
</tr>
<tr class="odd">
<td align="left">Quantity_2013</td>
<td align="left">Export or Import quantity in 2013</td>
</tr>
<tr class="odd">
<td align="left">Quantity_2014</td>
<td align="left">Export or Import quantity in 2014</td>
</tr>
<tr class="odd">
<td align="left">Quantity_2015</td>
<td align="left">Export or Import quantity in 2015</td>
</tr>
<tr class="odd">
<td align="left">Quantity_2016</td>
<td align="left">Export or Import quantity in 2016</td>
</tr>
<tr class="odd">
<td align="left">Quantity_2017</td>
<td align="left">Export or Import quantity in 2017</td>
</tr>  
</tbody>
</table>

<table>
<thead>
<tr class="header">
<th align="center" colspan="2">ExportValues & ImportValues</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="left">Country</td>
<td align="left">Trade Partner of Turkey</td>
</tr>
<tr class="odd">
<td align="left">Product_Code</td>
<td align="left">Traded Product Code in 4 digit</td>
</tr>
<tr class="odd">
<td align="left">Product_Label</td>
<td align="left">Product Definition</td>
</tr>  
<tr class="odd">
<td align="left">Value_2013</td>
<td align="left">Export or Import value in 2013</td>
</tr>
<tr class="odd">
<td align="left">Value_2014</td>
<td align="left">Export or Import value in 2014</td>
</tr>
<tr class="odd">
<td align="left">Value_2015</td>
<td align="left">Export or Import value in 2015</td>
</tr>
<tr class="odd">
<td align="left">Value_2016</td>
<td align="left">Export or Import value in 2016</td>
</tr>
<tr class="odd">
<td align="left">Value_2017</td>
<td align="left">Export or Import value in 2017</td>
</tr>
</tbody>
</table>

### Data Preparation, Cleaning and Manipulation

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(shiny)
library(tidyverse)
library(dplyr)
library(readxl)
library(leaflet)
library(reshape2)
library(gridExtra)
library(lubridate)
library(plotrix)
```

In order to read these excel files, firstly we create a function named "getDataSet":

```{r, warning=FALSE}

PATH <-"C:/Users/emrek/Google Drive/BDA/503-EssentialsOfDataAnalytics/GitHub/gpj18-group_four/Raw Data"
setwd(PATH)

getDataSet <- function(pPattern){
  nm <- list.files(path = PATH, pattern = pPattern)  
  if (exists("resultDataset")){
    rm("resultDataset")
  }
  for (file in nm){
    if (!exists("resultDataset")){
      resultDataset <- read_excel(file, col_names=FALSE, skip=1)
    }
    else{
      temp_dataset <-read_excel(file, col_names=FALSE, skip=1)
      resultDataset<-rbind(resultDataset, temp_dataset)
      rm(temp_dataset)
    }

  }
  return(resultDataset)

}

import_quantity_dataset <- getDataSet('*.mport.*uant.*')
import_value_dataset <- getDataSet('*.mport.*alu.*')
export_quantity_dataset <- getDataSet('*.xport.*uant.*')
export_value_dataset <- getDataSet('*.xport.*alu.*')

colnames(import_quantity_dataset) <- c("Country","Product_Code","Product_Label","Quantity_2013","Unit1","Quantity_2014","Unit2","Quantity_2015","Unit3","Quantity_2016","Unit4","Quantity_2017","Unit5")
colnames(import_value_dataset) <- c("Country","Product_Code","Product_Label","Value_2013","Value_2014","Value_2015","Value_2016","Value_2017")
colnames(export_quantity_dataset) <- c("Country","Product_Code","Product_Label","Quantity_2013","Unit1","Quantity_2014","Unit2","Quantity_2015","Unit3","Quantity_2016","Unit4","Quantity_2017","Unit5")
colnames(export_value_dataset) <- c("Country","Product_Code","Product_Label","Value_2013","Value_2014","Value_2015","Value_2016","Value_2017")

import_quantity_dataset$Product_Code<- gsub("'", "", import_quantity_dataset$Product_Code)
import_value_dataset$Product_Code<- gsub("'", "",import_value_dataset$Product_Code)
export_quantity_dataset$Product_Code<- gsub("'", "",export_quantity_dataset$Product_Code)
export_value_dataset$Product_Code<- gsub("'", "",export_value_dataset$Product_Code)

```

And by using the getDataSet function, we merged the Turkey's export-quantity, export-value, import-quantity, import-value tables for 10 selected countries.

```{r}
glimpse(export_quantity_dataset)
glimpse(export_value_dataset)
glimpse(import_quantity_dataset)
glimpse(import_value_dataset)
```

We have 13 and 8 variables in quantity and value tables respectively both for export and import.

Now, it is time to join quantity and value data for both export and import files. While joining the tables, we should also clear the repetitive variable "Unit" for each year.

Also, we need to join product labels in 2 digit product codes to 4 digit product codes and labels.

```{r, warning=FALSE}
#read the ProductCodes excel
file = "C:/Users/emrek/Google Drive/BDA/503-EssentialsOfDataAnalytics/GitHub/gpj18-group_four/2digitproductlist.xlsx"
prdCode <- read_excel(file, col_names=FALSE, skip=1)
prdCode$X__1 <- gsub("'", "", prdCode$X__1)

# join the import values and quantities
import_jTable <- import_quantity_dataset %>% 
  mutate( Unit= coalesce(Unit1,Unit2 ,Unit3 , Unit4,Unit5)) %>%
  select( Country, Product_Code , Unit, Quantity_2013, Quantity_2014, Quantity_2015, Quantity_2016, Quantity_2017) %>% 
  full_join(import_value_dataset,by = c("Country","Product_Code"))  
# seperate the first 2 digits of 4 digit product code and join the 2digit product code excel 
import_jTable <- import_jTable %>% mutate(PcodeTwo=substr(Product_Code,1,2))
import_jTable <- left_join(import_jTable,prdCode,by =c("PcodeTwo"="X__1") )

# join the export values and quantities
export_jTable <- export_quantity_dataset %>% 
  mutate( Unit= coalesce(Unit1,Unit2 ,Unit3 , Unit4,Unit5)) %>%
  select( Country, Product_Code , Unit, Quantity_2013, Quantity_2014,Quantity_2015,Quantity_2016,Quantity_2017) %>% 
  full_join(export_value_dataset,by = c("Country","Product_Code"))  
# seperate the first 2 digits of 4 digit product code and join the 2digit product code excel
export_jTable <- export_jTable %>% mutate(PcodeTwo=substr(Product_Code,1,2))
export_jTable <- left_join(export_jTable,prdCode,by =c("PcodeTwo"="X__1") )

```

Now we have joined tables for import and export stats of Turkey.

```{r}
glimpse(import_jTable)
glimpse(export_jTable)
```

Both tables now have 16 variables and 12,223 rows.

But there is a row in Product_Code column as "TOTAL", we must clear it.

```{r}
export_jTable <- export_jTable %>% filter(!grepl("TOTAL", Product_Code))
import_jTable <- import_jTable %>% filter(!grepl("TOTAL", Product_Code))
```

We have 16 variables in these tables however quantity and value variables repeat for each year. So actually, it is better to organize the table that it will have variables "Year", "Quantity", "Value" instead of 10 columns.

```{r}
#create new objects for each years' quantity
q2013<-import_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2013) %>% mutate(year="2013") %>% rename("Quantity"=Quantity_2013)
q2014<-import_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2014) %>% mutate(year="2014") %>% rename("Quantity"=Quantity_2014)
q2015<-import_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2015) %>% mutate(year="2015") %>% rename("Quantity"=Quantity_2015)
q2016<-import_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2016) %>% mutate(year="2016") %>% rename("Quantity"=Quantity_2016) 
q2017<-import_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2017) %>% mutate(year="2017") %>% rename("Quantity"=Quantity_2017)

#bind the objects
importQuant <- bind_rows(q2013,q2014,q2015,q2016,q2017)

#create new objects for each years' values
v2013<-import_jTable %>% select(Country,Product_Code, Value_2013) %>% rename("Values"=Value_2013) %>% mutate(year="2013")
v2014<-import_jTable %>% select(Country,Product_Code, Value_2014) %>% rename("Values"=Value_2014) %>% mutate(year="2014")
v2015<-import_jTable %>% select(Country,Product_Code, Value_2015) %>% rename("Values"=Value_2015) %>% mutate(year="2015")
v2016<-import_jTable %>% select(Country,Product_Code, Value_2016) %>% rename("Values"=Value_2016) %>% mutate(year="2016") 
v2017<-import_jTable %>% select(Country,Product_Code, Value_2017) %>% rename("Values"=Value_2017) %>% mutate(year="2017")

#bind the objects
importValues <- bind_rows(v2013,v2014,v2015,v2016,v2017)

#join binded quantity and value
import_jTableMELTED <- importQuant %>% left_join(importValues, by=c("Country","Product_Code","year"))

#Repeat with export table

q2013<-export_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2013) %>% mutate(year="2013") %>% rename("Quantity"=Quantity_2013)
q2014<-export_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2014) %>% mutate(year="2014") %>% rename("Quantity"=Quantity_2014)
q2015<-export_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2015) %>% mutate(year="2015") %>% rename("Quantity"=Quantity_2015)
q2016<-export_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2016) %>% mutate(year="2016") %>% rename("Quantity"=Quantity_2016) 
q2017<-export_jTable %>% select(Country, PcodeTwo, X__2 ,Product_Code, Product_Label, Unit, Quantity_2017) %>% mutate(year="2017") %>% rename("Quantity"=Quantity_2017)

exportQuant <- bind_rows(q2013,q2014,q2015,q2016,q2017)

v2013<-export_jTable %>% select(Country,Product_Code, Value_2013) %>% rename("Values"=Value_2013) %>% mutate(year="2013")
v2014<-export_jTable %>% select(Country,Product_Code, Value_2014) %>% rename("Values"=Value_2014) %>% mutate(year="2014")
v2015<-export_jTable %>% select(Country,Product_Code, Value_2015) %>% rename("Values"=Value_2015) %>% mutate(year="2015")
v2016<-export_jTable %>% select(Country,Product_Code, Value_2016) %>% rename("Values"=Value_2016) %>% mutate(year="2016") 
v2017<-export_jTable %>% select(Country,Product_Code, Value_2017) %>% rename("Values"=Value_2017) %>% mutate(year="2017")

exportValues <- bind_rows(v2013,v2014,v2015,v2016,v2017)

export_jTableMELTED <- exportQuant %>% left_join(exportValues, by=c("Country","Product_Code","year"))
```

Now, we have more useful tables, with 9 variables and 61100 rows.

```{r}
glimpse(export_jTableMELTED)
glimpse(import_jTableMELTED)
```

### Analysis

#### Breakdown of Total Export and Import of Turkey Over Countries

```{r}
ttable1=export_jTableMELTED%>%group_by(Country)%>%
  summarise(co_total=sum(Values,na.rm = TRUE))%>%
  arrange(desc(co_total,year))

exptable1=export_jTableMELTED%>%group_by(Country)%>%
  summarise(co_total=sum(Values,na.rm = TRUE))%>%
  arrange(desc(co_total,year))

pex1_table<-ttable1%>%
  mutate(perc_count=(co_total/sum(ttable1$co_total))*100)%>%
  select(Country,co_total,perc_count)

imptable1=import_jTableMELTED%>%group_by(Country)%>%
  summarise(co_total=sum(Values,na.rm = TRUE))%>%
  arrange(desc(co_total,year))

pim1_table<-imptable1%>%
  mutate(perc_count=(co_total/sum(ttable1$co_total))*100)%>%
  select(Country,co_total,perc_count)

lbls_i <- paste(pim1_table$Country,round(pim1_table$perc_count,1)) # add percents to labels
lbls_i <- paste(lbls_i,"%",sep="") # ad % to labels

pie3D(pim1_table$perc_count,labels=lbls_i,main = "Total Percentage Import Per Countries", explode=0.1, radius=.9,
      labelcex = 1.2, start=0.7)

lbls_e <- paste(pex1_table$Country,round(pex1_table$perc_count,1)) # add percents to labels
lbls_e <- paste(lbls_e,"%",sep="") # ad % to labels

pie3D(pex1_table$perc_count,labels=lbls_e,main = "Total Percentage Export Per Countries", explode=0.1, radius=.9,
             labelcex = 1.2, start=0.7)

```

At import side, Turkeys' biggest trade partners are China, Germany, Russia and US while they are Germany, UK, Italy and USA at export side. Germany and USA are the 2 countries exist for both type of trade.

When we breakdown this info to years:

```{r}
# Export Values in $ 
ttable=export_jTableMELTED%>%group_by(Country,year)%>%
  summarise(co_total=sum(Values,na.rm = TRUE))%>%
  arrange(desc(co_total,year))
#Import Values in $
vtable=import_jTableMELTED%>%group_by(Country,year)%>%summarise(co_total=sum(Values,na.rm = TRUE))%>%arrange(desc(co_total,year))

#Import & Export percentage of Countries

a<-ggplot(data=vtable, aes(x=Country, y=co_total,fill=year)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black")+
  ggtitle("Total Import Values in $ to Selected Countries")

b<-ggplot(data=ttable, aes(x=Country, y=co_total,fill=year)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black")+ggtitle("Total Export Values in $ to Selected Countries")

grid.arrange(a,b,  nrow = 2)
```

Except Russia, fluctuates over years are not seem so significant.

In order to deeply understand the data; we build up shiny app with Country, Product and Year variables to review export and import values.

Please follow the link for interactive graph: [Link](https://groupfour.shinyapps.io/tradeOrg)

##### YORUM VE IMAGE EKLENECEK

#### Analysis of Year by Year Change in Trade Values

```{r}
tableExport <- export_jTableMELTED %>% 
  group_by(Country, year) %>%
  summarise(totalExpV=sum(Values,na.rm=TRUE)) %>%
  arrange(Country, year) %>%
  mutate(ExpValChng = (totalExpV - lag(totalExpV))/lag(totalExpV)*100)

tableImport <- import_jTableMELTED %>% 
  group_by(Country, year) %>%
  summarise(totalImpV=sum(Values,na.rm=TRUE)) %>%
  arrange(Country, year) %>%
  mutate(ImpValChng = (totalImpV - lag(totalImpV))/lag(totalImpV)*100)

table <- full_join(tableExport,tableImport) %>%
  select(Country, year, Export="ExpValChng", Import = "ImpValChng") %>%
  melt(id=c("Country", "year")) %>%
  filter(year!=2013)
```

```{r, }
ggplot(table, aes(x=year)) +
  geom_point(aes(y=value, color=Country, shape=variable), size=4) + 
  facet_grid(~variable) +
   labs(title="YoY Total Value Change",
    x="Year",
    y="% Change in  Value",
    shape=c("Trade Type"))
```

```{r}

# Canada and Russia significantly deviates. Investigate these countries in detail with traded products

# First Canada

# Export Analysis

summarytableCanadaEx <- export_jTableMELTED %>%
  filter(Country=="Canada") %>%
  group_by(X__2, year) %>%
  summarise(ExportValue=sum(Values)) %>%
  arrange(X__2, year) %>%
  mutate(ExpPercChngYoY = (ExportValue - lag(ExportValue))/lag(ExportValue)*100) %>%
  mutate(ExpChngYoy=((ExportValue - lag(ExportValue)))) %>%
  filter(year!=2013)

mean(summarytableCanadaEx$ExpChngYoy)
median(summarytableCanadaEx$ExpChngYoy)
sd(summarytableCanadaEx$ExpChngYoy)
quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.25, 0.50, 0.75, 1))
IQR= quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.75)) - quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.25))
altsinir=quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.25)) - (1.5*IQR)
ustsinir=quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.75)) + (1.5*IQR)

summarytableCanadaEx <- summarytableCanadaEx %>%
  filter(abs(ExpChngYoy)>max(abs(altsinir),ustsinir))

plotCanEx <- ggplot(summarytableCanadaEx, aes(x=substr(X__2, start = 1, stop = 30))) +
  geom_point(aes(y=ExpChngYoy, color=year), size=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid (~year) +
  facet_grid(rows = vars(year)) +
  labs(title="YoY Total Value Change with Canada - Export",
       x="Product Group",
       y="Change in Export Value",
       Color=c("Year"))
```

```{r, echo=FALSE}
# Import Analysis

summarytableCanadaIm <- import_jTableMELTED %>%
  filter(Country=="Canada") %>%
  group_by(X__2, year) %>%
  summarise(ImportValue=sum(Values)) %>%
  arrange(X__2, year) %>%
  mutate(ImpPercChngYoY = (ImportValue - lag(ImportValue))/lag(ImportValue)*100) %>%
  mutate(ImpChngYoy=((ImportValue - lag(ImportValue)))) %>%
  filter(year!=2013)

mean(summarytableCanadaIm$ImpChngYoy)
median(summarytableCanadaIm$ImpChngYoy)
sd(summarytableCanadaIm$ImpChngYoy)
quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.25, 0.50, 0.75, 1))
IQR= quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.75)) - quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.25))
altsinir=quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.25)) - (1.5*IQR)
ustsinir=quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.75)) + (1.5*IQR)

summarytableCanadaIm <- summarytableCanadaIm %>%
  filter(abs(ImpChngYoy)>max(abs(altsinir),ustsinir))

plotCanIm <- ggplot(summarytableCanadaIm, aes(x=substr(X__2, start = 1, stop = 30))) +
  geom_point(aes(y=ImpChngYoy, color=year), size=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid (~year) +
  facet_grid(rows = vars(year)) +
  labs(title="YoY Total Value Change Canada - Import",
       x="Product Group",
       y="Change in Import Value",
       Color=c("Year"))
```

```{r, echo=FALSE}

# Second Russia

# Export Analysis

summarytableRussianEx <- export_jTableMELTED %>%
  filter(Country=="Russian") %>%
  group_by(X__2, year) %>%
  summarise(ExportValue=sum(Values)) %>%
  arrange(X__2, year) %>%
  mutate(ExpPercChngYoY = (ExportValue - lag(ExportValue))/lag(ExportValue)*100) %>%
  mutate(ExpChngYoy=((ExportValue - lag(ExportValue)))) %>%
  filter(year!=2013)

mean(summarytableRussianEx$ExpChngYoy)
median(summarytableRussianEx$ExpChngYoy)
sd(summarytableRussianEx$ExpChngYoy)
quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.25, 0.50, 0.75, 1))
IQR= quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.75)) - quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.25))
altsinir=quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.25)) - (1.5*IQR)
ustsinir=quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.75)) + (1.5*IQR)

summarytableRussianEx <- summarytableRussianEx %>%
  filter(abs(ExpChngYoy)>max(abs(altsinir),ustsinir))

plotRusEx <- ggplot(summarytableRussianEx, aes(x=substr(X__2, start = 1, stop = 30))) +
  geom_point(aes(y=ExpChngYoy, color=year), size=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid (~year) +
  facet_grid(rows = vars(year)) +
  labs(title="YoY Total Value Change Russia - Export",
       x="Product Group",
       y="Change in Export Value",
       Color=c("Year"))

# Import Analysis

summarytableRussianIm <- import_jTableMELTED %>%
  filter(Country=="Russian") %>%
  group_by(X__2, year) %>%
  summarise(ImportValue=sum(Values)) %>%
  arrange(X__2, year) %>%
  mutate(ImpPercChngYoY = (ImportValue - lag(ImportValue))/lag(ImportValue)*100) %>%
  mutate(ImpChngYoy=((ImportValue - lag(ImportValue)))) %>%
  filter(year!=2013)

mean(summarytableRussianIm$ImpChngYoy)
median(summarytableRussianIm$ImpChngYoy)
sd(summarytableRussianIm$ImpChngYoy)
quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.25, 0.50, 0.75, 1))
IQR= quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.75)) - quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.25))
altsinir=quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.25)) - (1.5*IQR)
ustsinir=quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.75)) + (1.5*IQR)

summarytableRussianIm <- summarytableRussianIm %>%
  filter(abs(ImpChngYoy)>max(abs(altsinir),ustsinir))

plotRusIm <- ggplot(summarytableRussianIm, aes(x=substr(X__2, start = 1, stop = 30))) +
  geom_point(aes(y=ImpChngYoy, color=year), size=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid (~year) +
  facet_grid(rows = vars(year)) +
  labs(title="YoY Total Value Change Russia - Import",
       x="Product Group",
       y="Change in Import Value",
       Color=c("Year")) 
```

```{r, fig.height = 8, fig.width = 16}

#install.packages("gridExtra")

library(gridExtra)

grid.arrange(plotCanEx, plotCanIm,  nrow = 1)
grid.arrange(plotRusEx, plotRusIm,  nrow = 1)

```

#### Analysis of Trade Deficit by Countries and Years

To discovery that country's effects to trade deficit, we calculated net_value which formulated as 'Export_value - Import_value'. And visualized that value on the basis of year.

First, group data;

```{r }
export_Product_Group <- export_jTableMELTED%>%
group_by(Country,PcodeTwo,X__2,year)%>%
summarize(Quantity = sum(Quantity), Values = sum(Values))%>%
rename(Product_Group_Code =PcodeTwo, Product_Group_Name = X__2,Export_Quantity = Quantity,Export_Value =Values)

import_Product_Group <- import_jTableMELTED%>%
group_by(Country,PcodeTwo,X__2,year)%>%
summarize(Quantity = sum(Quantity), Values = sum(Values))%>%
rename(Product_Group_Code =PcodeTwo, Product_Group_Name = X__2,Import_Quantity = Quantity,Import_Value =Values)

Export_Import_JTable <- export_Product_Group %>%
  full_join(import_Product_Group,by=c("Country","Product_Group_Code","Product_Group_Name","year")) %>%
mutate(Import_Quantity = replace(Import_Quantity, is.na(Import_Quantity),0),
        Export_Quantity = replace(Export_Quantity, is.na(Export_Quantity),0))%>%
select(Country,Product_Group_Code,Product_Group_Name,year,Export_Quantity,Import_Quantity,Export_Value,Import_Value)
```

Continue with visulation;

```{r}
####Net Value By Country###
Net_Value_By_Country <- Export_Import_JTable %>%
group_by(Country,year)%>%
summarize (Export_Value = sum(Export_Value), Import_Value = sum(Import_Value), 
           Net_Value = sum(Export_Value - Import_Value),Net_value_K = round((sum(Export_Value - Import_Value)/1000),digits=0))

# Total Net Value fiiled with Country on the basis of Year #
ggplot(Net_Value_By_Country, aes(x=year, y=Net_value_K)) + 
  geom_bar(stat='identity', aes(fill=Country), width=.6,position = position_stack(reverse = TRUE))  +
  labs(subtitle="Total Net Value fiiled with Country on the basis of Year ##", title= "Total Net Values") + 
  coord_flip()+
  theme(legend.position = "top")
```

As seen in the graph, only UK gives trade surplus for all years analysed and China, Russia, Germany and USA have the most significant effect on trade deficit.

##### An Alternative perspective to explore effects of countries

The effects of countries are changing year by year. So that as an alternative perspective we could average net value for all countries in five years and normalized that value to show which country's effects are more than others.

```{r }
#Normalized of Average Values for five years
Normalized_Values_By_Country <- Export_Import_JTable %>%
  #filter (year == 2017)%>%
  group_by(Country)%>%
  summarize (Net_Value = mean(Export_Value - Import_Value))%>%
  mutate(normalized_Net_value =  round((Net_Value  - mean(Net_Value))/sd(Net_Value ),2)) %>%
  mutate(value_Net_type =  ifelse(normalized_Net_value < 0, "below", "above"))

#Diverging Bars for normalized_Export_value
ggplot(Normalized_Values_By_Country, aes(x=Country, y=normalized_Net_value , label=normalized_Net_value)) + 
  geom_bar(stat='identity', aes(fill=value_Net_type), width=.5)  +
  scale_fill_manual(name="Normalised Value", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="seagreen1", "below"="Red")) +  
  labs(subtitle="Normalised Net Value for 5 years by Country'", 
       title= "Diverging Bars") + 
  coord_flip()
```

Considering the 5 years period, it is clear that China, Germany and Russia has contribution to trade deficit above average.

We drilled down the analysis for the countries which decreases the average, namely China, Russia and Germany. For this aim, we build up shiny app to review the products by defining the Country and ratio showing cumulative percentage of share of products over total import value of this country.

Please follow the link for interactive graph: [Link](https://groupfour.shinyapps.io/tradeOrgProAnly/)

Through the shiny app, it is understood that
.70% of import value from China is coming from 8 products 
.50% is from electrical machinery, equipments and machinery, mechanical appliances for China
.%70 is from commodities and mineral fuels for import from Russia
.only commodities constitute 43% of total import of Russia
.vehicles other than railway is the top product imported from Germany
.machinery, mechanical appliances are the second product imported from Germany and its share is so close to vehicles

##### IMAGE EKLENECEK

#### Correlation of Trade Deficit with FX Rate and BIST100 index

In order to understand the correlation between trade deficit/surplus and BIST100 index and USD/TL rates, we added this economic indicators to our analysis:

Read the USD and BIST100 index values and rename the columns

```{r}
fileUSD= "C:/Users/emrek/Google Drive/BDA/503-EssentialsOfDataAnalytics/GitHub/gpj18-group_four/EVDS.xlsx"
usdR <- read_excel(fileUSD, col_names=FALSE, skip=1)

colnames(usdR) <- c("DateR","ExchangeRate")
usd <- na.omit(usdR)
remove(usdR)

usdN<-usd %>% mutate(Date=dmy(DateR))
usdM<- usdN %>% group_by(year=year(Date)) %>% summarise(Rate=mean(ExchangeRate)*10000000)

fileBIST= "C:/Users/emrek/Google Drive/BDA/503-EssentialsOfDataAnalytics/GitHub/gpj18-group_four/BIST100.xlsx"
bist100R <- read_excel(fileBIST, col_names=FALSE, skip=1)

colnames(bist100R) <- c("Date",	"Year","Value",	"Açılış",	"Yüksek",	"Düşük",	"Hac.",	"Fark %")
bist100 <- bist100R  %>% select(Date,Year,Value) %>% 
  group_by(Year) %>% 
  summarise(Value = mean(Value)*1000)
```

Draw the graph to represent USD/TL rate, BIST100 index and Trade Deficit in time series.

```{r}
sumImport<- import_jTable %>% group_by(Country) %>% summarise(Value_2013=sum(Value_2013),Value_2014=sum(Value_2014),Value_2015=sum(Value_2015),Value_2016=sum(Value_2016),Value_2017=sum(Value_2017))
sumExport<- export_jTable %>% group_by(Country) %>% summarise(Value_2013=sum(Value_2013),Value_2014=sum(Value_2014),Value_2015=sum(Value_2015),Value_2016=sum(Value_2016),Value_2017=sum(Value_2017))
joinedTableDeficit<-sumExport %>%  inner_join(sumImport,by="Country") %>%
  mutate(Value_2013=Value_2013.y-Value_2013.x,Value_2014=Value_2014.y-Value_2014.x, Value_2015=Value_2015.y-Value_2015.x, Value_2016=Value_2016.y-Value_2016.x,
         Value_2017=Value_2017.y-Value_2017.x) %>%
  select(Country, Value_2013, Value_2014, Value_2015, Value_2016, Value_2017)

joinedTableDeficitMelt <- melt(joinedTableDeficit, id = c("Country"))
deficitPerCountry<-joinedTableDeficitMelt %>% mutate(year = case_when( variable == "Value_2013" ~ 2013,
                                                                       variable == "Value_2014" ~ 2014,
                                                                       variable == "Value_2015" ~ 2015,
                                                                       variable == "Value_2016" ~ 2016,
                                                                       variable == "Value_2017" ~ 2017)) %>%
  select(Country,year,value)

deficitPerYear <- deficitPerCountry %>%
  group_by(year) %>% summarise(value=sum(value))

p<-ggplot() + 
  geom_line(data=usdM,aes(x=year, y=Rate,group = 1, colour="USD_TL_Rate"))+
  geom_line(data=deficitPerYear,aes(x=year, y=value,group = 1, colour="Trade_Deficit"))+
  geom_line(data=bist100,aes(x=Year, y=Value,group = 1,colour="BIST100_Index")) +
  scale_colour_manual(name="Indicators",values=c(USD_TL_Rate="red", Trade_Deficit="blue", BIST100_Index="purple"))

print(p)

```

Trade deficit has negatively correlated with USD/TL rate between 2013 and 2016 as expected and for the year 2017 we saw the effect of stagflation - while the USD/TL rate increasing (accordingly we know that inflation rate is strongly positively correlatied with the inflation rate (inf rate: 2015-12:5.71%, 2016-12:9.94%, 2017-12:15.47%)) and weak or negative economic growth. 
Besides, there is not a direct correlation between BIST100 index and Trade Deficit Values.